---
title: "code-Red_module5"
author: "code-Red"
date: "2022-10-04"
output: html_document
bibliography: BIOL2140.bib
nocite: '@*'
---

```{r,include=FALSE}
library(tidyverse) 
library(knitr) 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The purpose of this project was to determine whether human body temperature varies according to activity level within individuals. A study done by @wunder1974effect revealed that the body temperature of Ord's kangaroo rat was higher during times of activity than inactivity. In order to determine whether a similar relationship exists in humans, temperature was recorded in 4 subjects during 5 minutes of rest (sitting on a couch) and 5 minutes of activity (jumping jacks). The results were plotted in order to determine whether there is a significant difference in human body temperature when active versus inactive.

# Methods

An Arduino DHT11 temperature sensor was set up following the instruction and file provided by @ada_2012 and was further modified to measure body temperature over time. The sensor was placed in the armpit of each group member to measure their axillary body temperature (°C) during periods of activity and inactivity. Body temperature was recorded 100 seconds after sitting on the couch and after five minutes of jumping jacks. The collected data was then plotted using a boxplot model in order to compare the active and at rest body temperatures across all the group members.

```{r, "Join Data into a Table", message=FALSE}

dat.f <- list.files(pattern = "couch|active.csv")

dat.l <- list() 
for(i in dat.f){
  met.dat<- unlist(strsplit(i,"_")) 
  who <- met.dat[1] 
  activity <- gsub(".csv","",met.dat[2])
  dat.l[[i]]<- read_csv(i)%>%
    mutate(who=who,activity=activity) 
}

dat <- do.call(rbind,dat.l) #combine the list into table
dat
```


# Results
```{r, "Plot Data"}
dat%>%
  ggplot()+geom_boxplot(aes(x=activity,y=degC))+facet_grid(.~who)+
  labs(caption = "Figure 1: Mean Body Temperature Between Activity Type By Subject") +
  theme(plot.caption = element_text(hjust=0.5))
```

The mean axial temperature between activity levels differed significantly in every subject, in that a decrease in temperature was observed after 5 minutes of jumping-jacks comparing to resting. This is indicated by the mean temperature between activity levels consistently falling out of each others' ±1 standard deviation within each subject.


# Discussion

As endotherms, humans maintain a constant body temperature independent of the environment, but slight variations as a result of activity are not unheard of. @hoffmann201224 found a difference of more than 1°C between peak measurement right after a 20min exercise and minimum sleep temperature within 24 hours. However, the mean body temperatures from our data showed significantly higher values in the resting condition than the active condition, which was unexpected and incongruent with @hoffmann201224. This could indicate systemic errors during data collection. Alternatively, axial temperature readings were taking under the armpit through clothing fabric for sanitary purposes, which could have resulted in cooling due to constant movement and slower heat transduction during the start of each trial. To further improve our study we could record temperature in contact with the body during the time of the exercise and compare it to the rest temperature to observe if the temperature would equalize due to our endothermic physiology.

# Author Contributions
Justin Lau - Arduino code, recording data, methods, plotting data\
Tingwei Hu - discussion & bib, edited Rmd setup, plot caption, Arduino setup (w/ Justin)\
Jinseo Kim - bib, methods revision, discussion revision\
Liz Asprinio - Introduction, methods, bibliography

# References