---
title: "code-Red_module5"
author: "code-Red"
date: "2022-10-04"
output: html_document
bibliography: BIOL2140.bib
nocite: '@*'
---

```{r,include=FALSE}
library(tidyverse) 
library(knitr) 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The purpose of this project was to determine whether human body temperature varies according to activity level. A study done by @wunder1974effect revealed that the body temperature of Ord's kangaroo rat was higher during times of activity than inactivity. In order to determine whether a similar relationship exists in humans, temperature was recorded in 4 subjects during 5 minutes of rest (sitting on a couch) and 5 minutes of activity (jumping jacks). The results were plotted in order to determine whether there is a significant difference in human body temperature when active versus inactive.

# Methods

An Arduino DHT__ temperature sensor was placed in the armpits of each group member to measure their axillary body temperature (°C). For each group member, body temperature was recorded for 100 seconds after sitting on the couch and five minutes of jumping jacks. 

```{r, "Join Data into a Table"}

dat.f <- list.files(pattern = "couch|active.csv") #find the files

dat.l <- list() #make an empty list

for(i in dat.f){
  met.dat<- unlist(strsplit(i,"_")) #split of the filename string
  who <- met.dat[1] #store the first element
  activity <- gsub(".csv","",met.dat[2]) #store second element, dropping CSV with gsub()
  dat.l[[i]]<- read_csv(i)%>%
    mutate(who=who,activity=activity) #read, then store data with additional columns of metada in list as position unique to i
  
}
dat <- do.call(rbind,dat.l) #combine the list into table
view(dat)

```


# Results
```{r, "Plot Data"}
dat%>%
  ggplot()+geom_boxplot(aes(x=activity,y=degC))+facet_grid(.~who)+
  labs(caption = "Mean Body Temperature Between Activity Type By Subject") +
  theme(plot.caption = element_text(hjust=0.5))
```

# Discussion
As endotherms, humans maintain a constant body temperature independent of the environment, but slight variations as a result of activity is not unheard of. @hoffman201224 found a difference of more than 1°C between peak measurement right after a 20min exercise and minimum sleep temperature within 24 hours. The lack of difference in mean body temperature between activity levels reflected by our data is incongruent with their findings. However, representative data from @hoffman201224 also showed that body temperature returns to a near-original level in 30 minutes after exercising. This suggests a time-sensitive nature of activity-dependent body temperature variation. In our study, the absence of substantial differences in body temperature recordings between activity levels could be caused by a mixture of insufficient activity intensity and success in heat dissipation. Further efforts could expand the scope of this study by taking metabolic rates and chronologically relevant nutrients intake into account.

# Author Contributions
Justin Lau - Arduino code, recording data, methods, plotting data
Tingwei Hu - discussion & bib, edited Rmd setup, plot caption

# References