---
title: "CodeRed_Module7"
author: "Liz Asprinio, Tingwei Hu, James Kim, Justin Lau"
date: "2022-11-27"
output: html_document
bibliography: BIOL3140_m7.bib
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, packages, include=FALSE}
library(tidyverse)
library(knitr)
library(vroom)
library(MuMIn)
```

```{r, Read files and data tidying, include=FALSE}


```

# Introduction

The muscle force-length relationship is governed by several factors. One aspect is the number of crossbridges formed in the regions of actin-myosin overlap of a sarcomere, which is directly related to the maximum amount of isometric force that can be produced by the muscle. Another aspect is mechanical advantage, the ratio between input (L_i) and output (L_o) force:
				$$MA = L_i/L_o$$
Greater mechanical advantage is described by a smaller ratio, or moving a load using a relatively small force. In the human body, this concept is illustrated by bones and joints acting as levers and pivots to minimize the force required by muscles to move a load. The arm is a lever and has a mechanical advantage of 0.2. However, the insertion angle of the muscles changes as the forearm flexes, which also contributes to output force. This complicates the study of force-length relationships, but research has shown that output force is sufficient to determine force-length relationships in input muscles (@rassier1999length).

# Methods





# Results




# Discussion



# Author Contribution




# References