---
title: "code-Red_module5"
author: "code-Red"
date: "2022-10-04"
output: html_document
bibliography: BIOL2140.bib
nocite: '@*'
---

```{r,include=FALSE}
library(tidyverse) #always
library(ape) #for evo analysis
library(nlme) #ditto
library(MuMIn) #for AIC
library(geiger) #for eco analysis
library(knitr) #for knitting options
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The purpose of this project was to determine whether human body temperature varies according to activity level. A study done by @wunder1974effect revealed that the body temperature of Ord's kangaroo rat was higher during times of activity than inactivity. In order to determine whether a similar relationship exists in humans, temperature was recorded in 4 subjects during 5 minutes of rest (sitting on a couch) and 5 minutes of activity (jumping jacks). The results were plotted in order to determine whether there is a significant difference in human body temperature when active versus inactive.

# Methods

An Arduino DHT__ temperature sensor was placed in the armpits of each group member to measure their axillary body temperature (°C). For each group member, body temperature was recorded for 100 seconds after sitting on the couch and five minutes of jumping jacks. 

```{r, "Join Data into a Table"}

dat.f <- list.files(pattern = "couch|active.csv") #find the files

dat.l <- list() #make an empty list

for(i in dat.f){
  met.dat<- unlist(strsplit(i,"_")) #split of the filename string
  who <- met.dat[1] #store the first element
  activity <- gsub(".csv","",met.dat[2]) #store second element, dropping CSV with gsub()
  dat.l[[i]]<- read_csv(i)%>%
    mutate(who=who,activity=activity) #read, then store data with additional columns of metada in list as position unique to i
  
}
dat <- do.call(rbind,dat.l) #combine the list into table
view(dat)

```


# Results
```{r, "Plot Data"}
dat%>%
  ggplot()+geom_boxplot(aes(x=activity,y=degC))+facet_grid(.~who) #plot to see what we get

```

# Discussion

# Author Contributions
Justin Lau - Arduino code, recording data, methods, plotting data


# References